---
title: "NHANES Part I"
author: "Tam Nguyen"
output: html_document
---

Quantitative Response, Quantitative Predictors: MULTIPLE LINEAR REGRESSION
-----------------------------------------

ORIGINAL SOURCE:  National Health and Nutrition Examination Survey, Centers for Disease Control and Prevention, http://www.cdc.gov/nchs/nhanes.htm.

Usually, body weight is determined by weighing on a scale.  But what if a scale is unavailable?  Could body weight be estimated using other, more easily measured, body measurements?  Suppose we only have a tape measure available, so we can only consider the body measurements obtained using a tape measure.  We would like to model weight as a linear (we hope!) function of length and circumference measurements, following "Occam’s razor"  or the "law of parsimony", which says "entities should not be multiplied beyond necessity" or "Of two equivalent theories or explanations, all other things being equal, the simpler one is to be preferred."

That is, we would like to have an "estimating equation" for body weight that provides the "best" estimate *with the fewest variables*.

We will be using data from NHANES 2009-2010 data to model body weight as a linear function of arm circumference and arm length.  We will restrict attention to only those individuals who are between 18 and 24 years old (inclusive).  That is, we will restrict attention to "young adults".

**Variables:**

Y:  body weight in kg

X1: upper arm circumference in cm

X2: upper arm length in cm

**Data set**: on Moodle as "NHANES-body.csv".  Download the file, then load it into the Console, then cut-and-paste the `read_csv` code into the chunk below.

Load data set (call it `nhanes`) and the mosiac package here:
```{r include=FALSE}
library(mosaic); library(readr)
nhanes <- read_csv("/Users/Tam/Documents/Junior1819/Fall2018/Stat/AppliedStat/ClassAct/Multi_linear/NHANES-body.csv") #replace this line with YOUR read_csv code
```
  
A. Investigating the Data
------------

1. You can read about the NHANES survey in the NHANES brochure (on the Moodle page).  From the information provided, do you feel the ever-present assumptions of representativeness and independence are satisfied with this data?  Why or why not?

- NHANES use complex survey designs that
oversample certain subpopulations like racial minorities. Naïve analysis of the original NHANES data can lead to mistaken conclusions. The percentages of people from each racial group in the data, for example,
are quite different from the way they are in the population.  


2. **Univariate Distributions:** Using graphical and descriptive statistics, describe the main features of the distribution of each variable, noting any unusual features.
```{r}
par(mfrow=c(1,2))
hist(nhanes$Upper.Arm.Length)
hist(nhanes$Arm.Circumference)
```

- The distribution looks odd at the beginning and the end of each the histogram. 

3. **Bivariate Relationships:**  We’d like to look at the relationship between all three of these variables.  Of course, we could just make three xyplots, but it might save a little time if we construct a scatterplot matrix.  This is a graph that makes scatterplots for all 

  a. Make the scatterplot matrix using the `pairs()` function.  To specify only these three variables (instead of all variables in the data set), use
```{r fig.height=5,fig.width=5}
pairs(~Weight+Upper.Arm.Length+Arm.Circumference, data=nhanes)
```

  b. For each pair of variables, describe the main features of the relationship, noting any unusual features.
- Weight with Arm.Circumference : all points are mostly compacted in a line, which we can see the linearily in the relationship
- Weight with Upper.Arm.Length: points are condense in once place, but has many outliner. 

B. Multiple Regression Model
---------------------------------

The simplest multiple linear regression model with two independent variables is:

$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon$, where $\epsilon_i$ are independent $N(0, \sigma_{\epsilon})$.

1. In general, what would the graph of this function look like?  (Hint: In 3 dimensions!)

- This graph presents in a three dimension plane. 

2. How many unknown parameters are in this model?  (Note that these parameters are the ones that will be estimated by the least-squares method.)
- There are three unknow parameters in this model.

3. The formulas for the least-squares estimates are similar, but more extensive, than those for the simple linear regression.  Of course, we will use R to calculate these estimates for us, using our old friend `lm()`.  The only change is we now indicate more than one predictor variable with a “+” sign in the formula argument:
```{r}
arm.lm <- lm(Weight~Arm.Circumference+Upper.Arm.Length, data=nhanes)
summary(arm.lm)
```

4. The fitted model is:
$Y = -80.06449 + 3.42556 Arm.Circumference + 1.29076 Upper.Arm.Length + \epsilon$, where $\epsilon_i$ are independent $N(0, \sigma_{\epsilon})$.

5. Interpret the estimates $\hat \beta_1$ and $\hat \beta_2$ in context.
- Arm.Circumference and Upper.Arm.Length are experessed in percentage (1-100)
- Estimate is 3.42556 for Arm.Circumference and 1.29076 for upper.arm.length.
- We estimate an expected 3.42 increase in arm circumference for every 1\% increase in percentage of weigth.
- We estimate an expected 1.29076 increase in upper arm length for every 1\% increase in percentage of weigth. 
- The t-test for \( H_0: \beta_{Agri} = 0 \) versus \( H_a: \beta_{Agri} \neq 0 \) is significant.

C. Diagnostics and Residual Analysis 
-----------------------------
Use the usual `plot(model)` to investigate the residual plots.  

1. Verify the validity of the model assumptions using the residual plots.
```{r}
par(mfrow=c(2,2))
plot(arm.lm)
```

- Normal Q - Q: Normal distribution
- Residuals vs fitted: too bunchy. it looks like it has non-constant variance. There maybe a curve in here, which is a problem.

2. What, if any, remedial measures (transformations, outlier corrections) are indicated by the residual analysis and diagnostics?

- We should try some transformations: 
- Residual (oy) / one var (ox) to see the relationship between residual values and Arm Circumferences var and Upper Arm Length var. 
```{r}
plot (arm.lm$residuals ~ arm.lm$model$Arm.Circumference)
plot (arm.lm$residuals ~ arm.lm$model$Upper.Arm.Length)
```

- According to graph above, the residuals vs the upper arm length values distributed constantly along the x-axis.
- The residuals vs the arm.circumference looks too bunchy and compact. 
3. Carry out any "fixes" you feel are necessary (but don’t add any new terms to the model yet!) and re-run the multiple linear regression model.  Verify the validity of the new model using residual plots.  Did you fix the problem(s) to your satisfaction?

- Maybe we should take the log(Weight) to expand the y-axis (weight).
```{r}
nhanes$trans_y = log(nhanes$Weight)
pairs(~trans_y+Upper.Arm.Length+Arm.Circumference, data=nhanes)
```

- We still can see the linear relationship between log(weight) vs arm.circumference and log(weight) vs upper.arm.length.


```{r}
new_arm.lm <- lm(trans_y~Arm.Circumference+Upper.Arm.Length, data=nhanes)
summary(new_arm.lm)
```
```{r}
par(mfrow=c(2,2))
plot(new_arm.lm)
```

- Normal Q-Q: Normal distribution
- Residuals vs fitted values: Looks good with constant variance  
- The new model has p-value ~ 0, which means this model is useful. Similarity, arm.circumference and upper.arm.length do have a linear relationship with the log(Weight).

D. Inferences About $\beta$ Parameters
-----------------------------------------

1. Explain why it is not appropriate in this setting to make an inference about $\beta_0$.

- In the formula, if all x-values equal to zero, y-value = $\beta_0$. It means that if a person that has upper.arm.length and arm.circumference both equal to 0, their weight will approximately equal to -80.0644, which doesn't make sense.
2. Construct a 90% confidence interval for $\beta_1$ and interpret the interval in context.
```{r}
confint(arm.lm, level = 0.95)
```
- The 90% confidence interval for $\beta_1$ is [3.316596 4.534532].
- Similarly, 90% of arm.circumference data is within the confidence interval.

3. Conduct a test of significance for $\beta_2$.
- We can get this restult by looking at the summary(arm.lm) above: p-value ~ 0, which means that the test for $\beta_2$ is significant. In another word, the upper.arm.length is a good variable to predict weight. 

4. Conduct the test of significance **for the entire model** using the ANOVA table.

```{r}
anova(arm.lm)
```
- F-statstic: 3265 on 2 and 795 DF
- In particular, we are testing correlation = 0 (the Null) and correlation \neq 0 (not the null). p-value < 2.2e-16 ~ 0, which means upper.arm.length and arm.circumference is useful in predicting weight. 

5. Interpret $R^2$ in context.
- $R^2$ = 0.8915
- 90 Percent of variablity of weight is explained by this model. 

